---
title: "lab10"
author: "Claire Madden"
date: "3/24/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      message = FALSE,
                      warning = FALSE)

library(tidyverse)
library(here)

library(DBI)
library(RSQLite)

```


### Connect to a database, and tell R we are using SQLite:
```{r}

#establish connection to the database
# drv = driver, dbname = database name
lter_db <- DBI::dbConnect(drv = RSQLite::SQLite(),
                          dbname = here("data", "marine.sqlite"))



```


### What tables exist in the database?
To check which tables exist (table names) in a database, use 'DBI::dbListTables()'

```{r}

DBI::dbListTables(lter_db)

```

### Which variables are in a database table?
Use 'DBI::dbListFields'

```{r}

DBI::dbListFields(lter_db, "fish_data")


```

```{r}

DBI::dbListFields(lter_db, "inverts_data")

```


```{r}

DBI::dbListFields(lter_db, "lobsters_data")

```

#### Read in a database table as an R dataframe!

Use 'DBI::dbReadTable()'

```{r}
fish_df <- DBI::dbReadTable(lter_db, "fish_data")
class(fish_df)
#now this dataframe is usable like any other with R stuff we already know
```

#### Query a database using SQL (need to use an SQL code chunk!)
- SELECT : gets columns
- FROM : what table in a database to get something from
- Use * to indicate "get all" 

```{sql connection=lter_db}

SELECT * FROM fish_data

```


To get a subset of the columns, we can just list them after 'SELECT', seperated by a comma:
if you want to connect to the same database in a bunch of different code chunks you can set as a global option using 'connection = (database name)'

```{sql connection=lter_db}

select year, site from fish_data

```


We might want to only query for ROWS (observations) that match a given condition we set:

- in R: dplyr::filter()
- in SQL: WHERE

```{sql connection=lter_db}

select * from fish_data where year == 2016

```


Now let's query to keep all columns from fish_data but only if the common_name variable matches "garibaldi":
```{sql connection=lter_db}
select common_name, site from fish_data where common_name == "garibaldir"
```

In dplyr: filter for an and statement: , or &
SQL: "and"

Example: Query to only keep observations where common_name == "garibaldi" AND site == "mohk"
```{sql connection=lter_db}
select * from fish_data where common_name == "garibaldi" and site == "mohk"
```


To write an "or" statement: 
in R: |, %in%
in SQL: "or"

Example: Query to keep observations where year is 2017 OR site is "abur"

```{sql connection=lter_db}
select * from fish_data where year == 2017 or site == "abur"
```

#### Store an output from SQL querying

add in the code chunk header 'output.var = "name_object"'

```{sql connection=lter_db, output.var = "fish_2017_df"}
select * from fish_data where year == 2017
```

#### Query using SQL from an R code chunk

Use 'DBI::dbSendQuery()'

```{r}
garibaldi <- dbSendQuery(lter_db, "SELECT * FROM fish_data WHERE common_name == 'garibaldi'")
# R can't do nested double quotes, which is why we used single quote to define common_name to match to
# query happened but it still doesn't know we want it as a dataframe

gar_df <- dbFetch(garibaldi)
```

### Query using dplyr-ish code

Use 'dbplyr'

```{r}
fish_tbl <- tbl(lter_db, "fish_data")

# some wrangling : 

yearly_fish <- fish_tbl %>% 
  group_by(year) %>% 
  summarize(
    total_fish = sum(total_count)
  )
# this is still recognized as a SQL type thing, but the table hasn't actually been created yet, if we want to see waht this would look like it SQL :

yearly_fish %>% show_query()
# this shows the SQL code that you would need to use to do the same thing that we just did in dplyr !!!~!~!!!!!

yearly_fish %>% collect()
```


#### Adding or removing tables from a SQLite database

Use 'DBI::dbWriteTable' and 'DBI::dbRemoveTable' functions to add or remove a table from an existing database (or can use these to create a new database)

```{r}
dbWriteTable(lter_db, "mtcars_df", mtcars)

# Check to see what tables are in the database now!

DBI::dbListTables(lter_db)

# now we're going to get rid of it

DBI::dbRemoveTable(lter_db, "mtcars_df")

DBI::dbListTables(lter_db)
```









